%%
%% This is file `marlowe-poemscol.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% dramatist.dtx  (with options: `pmarlowe')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from marlowe-poemscol.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file dramatist.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%% dramatist.dtx
%% Copyright (C) 2003-2005 Massimiliano Dominici
%% \CharacterTable%%  {Upper-case    \A\B\C\D\E\F\G\H\I\J\K\L\M\N\O\P\Q\R\S\T\U\V\W\X\Y\Z
%%   Lower-case    \a\b\c\d\e\f\g\h\i\j\k\l\m\n\o\p\q\r\s\t\u\v\w\x\y\z
%%   Digits        \0\1\2\3\4\5\6\7\8\9
%%   Exclamation   \!     Double quote  \"     Hash (number) \#
%%   Dollar        \$     Percent       \%     Ampersand     \&
%%   Acute accent  \'     Left paren    \(     Right paren   \)
%%   Asterisk      \*     Plus          \+     Comma         \,
%%   Minus         \-     Point         \.     Solidus       \/
%%   Colon         \:     Semicolon     \;     Less than     \<
%%   Equals        \=     Greater than  \>     Question mark \?
%%   Commercial at \@     Left bracket  \[     Backslash     \\
%%   Right bracket \]     Circumflex    \^     Underscore    \_
%%   Grave accent  \`     Left brace    \{     Vertical bar  \|
%%   Right brace   \}     Tilde         \~}
%%
%%\section{The GNU General Public License}
%% This file shows some basic features of package `dramatist', when
%% typesetting a drama in verse. Here we use package `poemscol' as a
%% support for text in verse; if you usually use package `verse'
%% instead, you may give a look to file `marlowe-verse.tex'.
%%
%% The source for this example is taken from Marlowe's `Jew of
%% Malta'. I took leave to insert division in scenes, which is not
%% present in the original edition, in order to show more features.

\documentclass{book}
\usepackage{fancyhdr,newmarn,geometry}
\usepackage{keyval,ifthen,newcropmark}
\usepackage{poemscol}

%% Load package `dramatist' after `poemscol'
\usepackage{dramatist}

%% comment out one of the following lines (and comment out the
%% previous one) if you want line numbering per act or per scene.
%%\usepackage[lnpa]{dramatist}
%%\usepackage[lnps]{dramatist}

%% We may change some parameters in the look of acts and scenes:
%%\renewcommand{\actnamefont}{\bfseries\Large}
%%\renewcommand{\theact}{\Roman{act}}
%%\renewcommand{\scenenamefont}{\bfseries\large}
%%\renewcommand{\thescene}{\Roman{scene}}

%% We may change some parameters in the look of characters:
%%\renewcommand{\castfont}{\bfseries}
%%\renewcommand{\speaksfont}{\itshape}
%%\renewcommand{\speaksdel}{:}

%% We may change some parameters in the look of stage directions:
%%\StageDirConf{\begin{center}\begin{minipage}{.4\textwidth}\bfseries}{\end{minipage}\end{center}}

%% Comment out the following lines if you want to insert a vertical
%% space between speeches.
%%\makeatletter
%%\renewcommand{\speakstab}{\if@stagedir\global\@stagedirfalse\else\bigskip\fi\hspace{\speaksskip}}
%%\makeatother

%% Maybe you want acts and scenes print their marks in the headings.
%% The following lines should work.
%%\renewcommand{\drampermark}[1]{\volumeheader{#1}}
%%\renewcommand{\actmark}[1]{\leftheader{#1}}
%%\renewcommand{\scenemark}[1]{\volumeheader{#1}}
%%\fancyhf{}
%%\fancyhead[CO]{\scshape \volumeheadervalue}
%%\fancyhead[CE]{\scshape \leftheadervalue}
%%\fancyfoot[CE]{\thepage}
%%\fancyfoot[CO]{\thepage}

\author{Christopher Marlowe}
\title{The Jew of Malta}
\date{}

\begin{document}
\begin{titlepage}
\maketitle
\end{titlepage}

\tableofcontents

%% We define some characters appearing in the play.
\Character[FERNEZE, governor of Malta.]{Ferneze}{fer}
\Character[LODOWICK, his son.]{Lodowick}{lod}
\Character[SELIM CALYMATH, son to the Grand Seignior.]{Calymath}{cal}
\Character[MARTIN DEL BOSCO, vice-admiral of Spain.]{Martin Del Bosco}{mar}
\Character[MATHIAS, a gentleman.]{Mathias}{mat}
%% group of characters.
\begin{CharacterGroup}{friars.}
\GCharacter{JACOMO,}{Jacomo}{jac}
\GCharacter{BARNARDINE,}{Barnardine}{barn}
\end{CharacterGroup}
\Character[BARABAS, a wealthy Jew.]{Barabas}{bar}
\Character[ITHAMORE, a slave.]{Ithamore}{ith}
\Character[PILIA-BORZA, a bully, attendant to BELLAMIRA.]{Pilia-Borza}{pb}
%% the following three collective characters appear in the play as
%% single instances, so we don't need define commands and entries
%% for them.
\Character[Two Merchants.]{}{}
\Character[Three Jews.]{}{}
\Character[Knights, Bassoes, Officers, Guard, Slaves, Messenger, and Carpenters]{}{}
\Character[KATHARINE, mother to MATHIAS.]{Katharine}{kat}
\Character[ABIGAIL, daughter to BARABAS.]{Abigail}{abi}
\Character[BELLAMIRA, a courtezan.]{Bellamira}{bel}
\Character[ABBESS.]{Abbess}{abb}
\Character[NUN.]{Nun}{nun}
%% This character doesn't appear in the dramatist personae list.
\Character{Machiavel}{mac}

%% We call the dramatis personae list.
\DramPer

%% The Prologue: we use \Act, but prevent it from printing \actname
%% and \theact.
\setcounter{secnumdepth}{-1}
\renewcommand{\printacttitle}[1]{\centering\acttitlefont #1}
%% This is needed only with customized headings.
%%\renewcommand{\actname}{}
%%\renewcommand{\theact}{}
\Act{Prologue}

\StageDir{Enter \mac.}

\begin{drama*}
\macspeaks
Albeit the world think Machiavel is dead,\verseline
Yet was his soul but flown beyond the Alps;\verseline
And, now the Guise is dead, is come from France,\verseline
To view this land, and frolic with his friends.\verseline
To some perhaps my name is odious;\verseline
But such as love me, guard me from their tongues,\verseline
And let them know that I am Machiavel,\verseline
And weigh not men, and therefore not men's words.\verseline
Admir'd I am of those that hate me most:\verseline
Though some speak openly against my books,\verseline
Yet will they read me, and thereby attain\verseline
To Peter's chair; and, when they cast me off,\verseline
Are poison'd by my climbing followers.\verseline
I count religion but a childish toy,\verseline
And hold there is no sin but ignorance.\verseline
Birds of the air will tell of murders past!\verseline
I am asham'd to hear such fooleries.\verseline
Many will talk of title to a crown:\verseline
What right had Caesar to the empery?\verseline
Might first made kings, and laws were then most sure\verseline
When, like the Draco's, they were writ in blood.\verseline
Hence comes it that a strong-built citadel\verseline
Commands much more than letters can import:\verseline
Which maxim had Phalaris observ'd,\verseline
H'ad never bellow'd, in a brazen bull,\verseline
Of great ones' envy:  o' the poor petty wights\verseline
Let me be envied and not pitied.\verseline
But whither am I bound?  I come not, I,\verseline
To read a lecture here in Britain,\verseline
But to present the tragedy of a Jew,\verseline
Who smiles to see how full his bags are cramm'd;\verseline
Which money was not got without my means.\verseline
I crave but this,--grace him as he deserves,\verseline
And let him not be entertain'd the worse\verseline
Because he favours me.\verseline
\direct{Exit.}
\end{drama*}

%% Now we restore the default settings.
\setcounter{act}{0}
\setcounter{secnumdepth}{0}
\renewcommand{\printacttitle}[1]{\acttitlefont #1}
%% This is needed only with customized headings.
%%\renewcommand{\actname}{Act}
%%\renewcommand{\theact}{\roman{act}}
%%\renewcommand{\theact}{\Roman{act}}

%% Without a title, the command to be used is \act. We add a
%% footnote, just to show a feature.
\act[\footnote{This is the first act.}]

\scene
%% Scene are not marked in the original edition. I have introduced
%% them (somehow arbitrarily, I suppose) to give a complete review
%% of the available features.

\StageDir{\bar discovered in his counting-house, with heaps
of gold before him.}

\begin{drama*}
\barspeaks So that of thus much that return was made;\verseline
And of the third part of the Persian ships\verseline
There was the venture summ'd and satisfied.\verseline
As for those Samnites, and the men of Uz,\verseline
That bought my Spanish oils and wines of Greece,\verseline
Here have I purs'd their paltry silverlings.\verseline
Fie, what a trouble 'tis to count this trash!\verseline
Well fare the Arabians, who so richly pay\verseline
The things they traffic for with wedge of gold,\verseline
Whereof a man may easily in a day\verseline
Tell that which may maintain him all his life.\verseline
The needy groom, that never finger'd groat,\verseline
Would make a miracle of thus much coin;\verseline
But he whose steel-barr'd coffers are cramm'd full,\verseline
And all his life-time hath been tired,\verseline
Wearying his fingers' ends with telling it,\verseline
Would in his age be loath to labour so,\verseline
And for a pound to sweat himself to death.\verseline
Give me the merchants of the Indian mines,\verseline
That trade in metal of the purest mould;\verseline
The wealthy Moor, that in the eastern rocks\verseline
Without control can pick his riches up,\verseline
And in his house heap pearl like pebble-stones,\verseline
Receive them free, and sell them by the weight;\verseline
Bags of fiery opals, sapphires, amethysts,\verseline
Jacinths, hard topaz, grass-green emeralds,\verseline
Beauteous rubies, sparkling diamonds,\verseline
And seld-seen costly stones of so great price,\verseline
As one of them, indifferently rated,\verseline
And of a carat of this quantity,\verseline
May serve, in peril of calamity,\verseline
To ransom great kings from captivity.\verseline
This is the ware wherein consists my wealth;\verseline
And thus methinks should men of judgment frame\verseline
Their means of traffic from the vulgar trade,\verseline
And, as their wealth increaseth, so inclose\verseline
Infinite riches in a little room.\verseline
But now how stands the wind?\verseline
Into what corner peers my halcyon's bill?\verseline
Ha! to the east? yes.  See how stand the vanes--\verseline
East and by south:  why, then, I hope my ships\verseline
I sent for Egypt and the bordering isles\verseline
Are gotten up by Nilus' winding banks;\verseline
Mine argosy from Alexandria,\verseline
Loaden with spice and silks, now under sail,\verseline
Are smoothly gliding down by Candy-shore\verseline
To Malta, through our Mediterranean sea.--\verseline
But who comes here?\verseline
%% We need another character
\Character{Merchant}{mer}
\direct{Enter a \mer.}
How now!\verseline

\merspeaks Barabas, thy ships are safe,\verseline
Riding in Malta-road; and all the merchants\verseline
With other merchandise are safe arriv'd,\verseline
And have sent me to know whether yourself\verseline
Will come and custom them.\verseline

\barspeaks The ships are safe thou say'st, and richly fraught?\verseline

\merspeaks They are.\verseline

\barspeaks Why, then, go bid them come ashore,\verseline
And bring with them their bills of entry:\verseline
I hope our credit in the custom-house\verseline
Will serve as well as I were present there.\verseline
Go send 'em threescore camels, thirty mules,\verseline
And twenty waggons, to bring up the ware.\verseline
But art thou master in a ship of mine,\verseline
And is thy credit not enough for that?\verseline
\end{drama*}

\begin{center}
[\dots]
\end{center}

\scene
%%Other characters.

\Character{knights}{knights}
\Character{officers}{officers}
\Character{bassoes}{bassoes}
\Character{First Basso}{fb}

%% An alternative formulation for \StageDir.
\begin{stagedir}
Enter \fer, governor of Malta, \knights, and \officers;
met by \cal, and \bassoes of the Turk.
\end{stagedir}

\begin{drama*}
\ferspeaks Now, bassoes, what demand you at our hands?\verseline

\fbspeaks Know, knights of Malta, that we came from Rhodes,\verseline
From Cyprus, Candy, and those other isles\verseline
That lie betwixt the Mediterranean seas.\verseline

\ferspeaks What's Cyprus, Candy, and those other isles\verseline
To us or Malta? what at our hands demand ye?\verseline

\calspeaks The ten years' tribute that remains unpaid.\verseline

\ferspeaks Alas, my lord, the sum is over-great!\verseline
I hope your highness will consider us.\verseline

\calspeaks I wish, grave governor, 'twere in my power\verseline
To favour you; but 'tis my father's cause,\verseline
Wherein I may not, nay, I dare not dally.\verseline
\end{drama*}

\begin{center}
[\dots]
\end{center}

\act

\scene

\StageDir{Enter \bar with a light.}

\begin{drama*}
\barspeaks Thus, like the sad-presaging raven, that tolls\verseline
The sick man's passport in her hollow beak,\verseline
And in the shadow of the silent night\verseline
Doth shake contagion from her sable wings,\verseline
Vex'd and tormented runs poor Barabas\verseline
With fatal curses towards these Christians.\verseline
The incertain pleasures of swift-footed time\verseline
Have ta'en their flight, and left me in despair;\verseline
And of my former riches rests no more\verseline
But bare remembrance; like a soldier's scar,\verseline
That has no further comfort for his maim.--\verseline
O Thou, that with a fiery pillar ledd'st\verseline
The sons of Israel through the dismal shades,\verseline
Light Abraham's offspring; and direct the hand\verseline
Of Abigail this night! or let the day\verseline
Turn to eternal darkness after this!--\verseline
No sleep can fasten on my watchful eyes,\verseline
Nor quiet enter my distemper'd thoughts,\verseline
Till I have answer of my Abigail.\verseline
\end{drama*}

\StageDir{Enter \abi above.}

\begin{drama*}
\abispeaks Now have I happily espied a time\verseline
To search the plank my father did appoint;\verseline
And here, behold, unseen, where I have found\verseline
The gold, the pearls, and jewels, which he hid.\verseline

\barspeaks Now I remember those old women's words,\verseline
Who in my wealth would tell me winter's tales,\verseline
And speak of spirits and ghosts that glide by night\verseline
About the place where treasure hath been hid:\verseline
And now methinks that I am one of those;\verseline
For, whilst I live, here lives my soul's sole hope,\verseline
And, when I die, here shall my spirit walk.\verseline
\end{drama*}

\begin{center}
[\dots]
\end{center}

\end{document}

\endinput
%%
%% End of file `marlowe-poemscol.tex'.
